---
title: "Project title"
subtitle: "STA/ISS 313 - Project 1"
author: "Team Name"
format: html
editor: visual
---

## Abstract

(1 paragraph): Project abstract.

------------------------------------------------------------------------

## Introduction

(1-2 paragraphs): Brief introduction to the dataset. You may repeat some of the information about the dataset provided in the introduction to the dataset on the TidyTuesday repository, paraphrasing on your own terms. Imagine that your project is a standalone document and the grader has no prior knowledge of the dataset.

## Question 1: Title that relates to the question you're answering

### Introduction

(1-2 paragraphs): Introduction to the question and what parts of the dataset are necessary to answer the question. Also discuss why you're interested in this question.

### Approach

(1-2 paragraphs): Describe what types of plots you are going to make to address your question. For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.

### Analysis

(2-3 code blocks, 2 figures, text/code comments as needed): In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides. You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so. All plots must be made with **ggplot2**. Do not use base R or lattice plotting functions.

```{r}
#| label: latitude and longitude data generation

# raw_contour <- read_delim(
#   "data/FM_service_contour_current.txt",
#   delim = "|"
# )
# 
# conv_contour <- raw_contour |>
#   select(-last_col()) |>
#   set_names(nm = c(
#     "application_id", "service", "lms_application_id", "dts_site_number", "transmitter_site",
#     glue::glue("deg_{0:360}")
#   ))

# lng_contour <- conv_contour |>
#   separate(
#     transmitter_site, 
#     into = c("site_lat", "site_long"), 
#     sep = " ,") |>
#   pivot_longer(
#     names_to = "angle",
#     values_to = "values",
#     cols = deg_0:deg_360
#   ) |>
#   mutate(
#     angle = str_remove(angle, "deg_"),
#     angle = as.integer(angle)
#   ) |>
#   separate(
#     values,
#     into = c("deg_lat", "deg_lng"),
#     sep = " ,"
#   )

# site_lat_long <- lng_contour %>% 
#   select(application_id, service, lms_application_id, site_lat, site_long)
# 
# site_lat_long <- unique(site_lat_long)
# 
# write_csv(site_lat_long, "data/site_lat_long.csv")

```

```{r}
#| label:  creating new genre column

library(tidyverse)

state_stations <- read_csv(file = 'data/state_stations.csv')
coords <- read_csv(file = 'data/site_lat_long.csv')
station_info <- read_csv(file='data/station_info.csv')

Genre_Lists <- list(

   c('Religious', 'Christian', 'Gospel', 'Religion', 'Catholic', 'worship'),
   c('Urban', 'R&B', 'Blues', 'Rhythmic','Rhythym and Blues','Hip Hop', 'mainstream urban', 'Urban adult contemporary', 'Urban contemporary'),
      c('Pop', 'Contemporary Hit', 'adult contemporary', 'Top 40 (CHR)', 'Hot AC', 'CHR'),
  c( 'Rock', 'Alternative','Indie'),
  
  c('Country','Southern', 'Bluegrass'),
  
  c('News', 'News/Talk', 'Talk'))

tmpstr <- Reduce(function(s, lst) gsub(paste0("(", paste(lst, collapse = "|"), ")"), lst[1], s, ignore.case = TRUE),
                 Genre_Lists, init = state_stations$format)
tmpGenres <- regmatches(tmpstr,
                        gregexpr(paste0("(", paste0(sapply(Genre_Lists, `[[`, 1), collapse = "|"), ")"),
                                 tmpstr, ignore.case = TRUE))
tmpGenres <- replace(tmpGenres, sapply(tmpGenres, Negate(length)), NA_character_)
tmpGenres <- sapply(tmpGenres, tail, n = 1)
tmpGenres[ is.na(tmpGenres) ] <- "Not Found"
state_stations$Genre <- tmpGenres

state_stations%>%
  group_by(Genre)%>%
  summarise(n())
```

```{r}
#| label: joining all DFs together

new <-state_stations |> dplyr::right_join(station_info, by = c("call_sign"))
new %>%
  group_by(state)%>%
  summarise(n())
```

```{r}
#| label: US choropleth map

library(maps)

us_states <- map_data("state")

ggplot() +
  geom_polygon(data = us_states,
               aes(x = long, y = lat, group = group),
               color = "black", fill = "lightpink")

number_stations <- new %>%
  group_by(state, Genre) %>%
  summarise(n = n())

number_stations <- drop_na(number_stations)
number_stations$state <- tolower(number_stations$state)

merged_stations <- full_join(us_states, number_stations, 
                             by = c("region" = "state"))

ggplot() +
  geom_polygon(data = merged_stations, 
               aes(x = long, y = lat, group = group, fill = n), 
               color = "black") 
  facet_wrap(~ Genre) +
    theme_none()



```

```{r}
map <- map_data("county")
nc <- subset(map, region =="north carolina")

ggplot(data = nc) + 
  geom_polygon(aes(x = long, y = lat, group = group), 
               color = "black", fill = "blue")

```

### Discussion

(1-3 paragraphs): In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots. Speculate about why the data looks the way it does.

## Question 2: Identifying correlation between population size and the number of radio stations in various states/regions of the United States

### Introduction

(1-2 paragraphs): Introduction to the question and what parts of the dataset are necessary to answer the question. Also discuss why you're interested in this question.

Population size and density varies across the United States. For example, according to *World Population Review* (<https://worldpopulationreview.com/state-rankings/state-densities>), in 2023 Wyoming had a population of 581,075 and a population density of about 6 people per square mile. Meanwhile, New York has a population of 19,300,000 people and a population density of about 410 people per square mile. Given this radical variability in population and population density, it may be possible to find a correlation between these quantities and the variability in the number of radio stations in each state. Exploring this relationship may help us understand how a population or population density influences the number of radio stations that exist in a particular state or region.

### Approach

(1-2 paragraphs): Describe what types of plots you are going to make to address your question. For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.

### Analysis

(2-3 code blocks, 2 figures, text/code comments as needed): In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides. You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so. All plots must be made with **ggplot2**. Do not use base R or lattice plotting functions.

### Discussion

(1-3 paragraphs): In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots. Speculate about why the data looks the way it does.
