---
title: "Stats FM"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
```

# Dataset

The data comes from a Wikipedia list of lists titled ​​List of Radio Stations in the United States (<https://en.wikipedia.org/wiki/Lists_of_radio_stations_in_the_United_States>). The list contains 50 links to radio stations categorized by state. Each of the 50 links lead to a table of FCC-licensed radio stations in the state. Here is a link to Alabama\'s (<https://en.wikipedia.org/wiki/List_of_radio_stations_in_Alabama>). The Wikipedi tables have 5 columns: Call Sign, Frequency, City of Liscence, Liscencee, and Format. Here is a description of the column names

-   Call sign: a unique identifier for a transmitter station

-   Frequency: number of cycles a radio wave is transmitted per second

-   City of License: the city where the station is officially licenseded to serve (https://en.wikipedia.org/wiki/City_of_license#:\~:text=In%20American%2C%20Canadian%2C%20and%20Mexican,by%20that%20country's%20broadcast%20regulator.)

-   Licensee: holder of the license

-   Format: general content broadcast by the station

The data collected from all 50 states were then aggregated to give the final data this project is using. 

The raw data, before wrangling, can be found in FCC\'s predicted service contour data points which is released regularly approximately at 10:00 AM Eastern Time daily (<https://www.fcc.gov/media/radio/fm-service-contour-data-points>). The details of how to read the data and its formats are given on the website. In addition, FCC\'s FM Query Broadcast Station Search can be used to look up a list of licensed stations (<https://www.fcc.gov/media/radio/fm-query>).

The data that is loaded from tidytuesday 2022-11-08 contains two datasets - state_stations and station_info. State_station contains 17,186 rows of the above data with 6 rows - the above 5 along with a state column that has states as value. Station_info contains 2,065 rows of station information data with 6 columns - call_sign, facility_id, service, licensee, status, and details. Here are the details for the columns

-   call_sign: a unique identifier for a transmitter station

```{=html}
<!-- -->
```
-   facility_id: facility id, 2,065 unique ids

-   service: frequency (FM)

-   licensee: holder of the license

-   status: status of the station (licensed, licensce ccanceled, licensced and silent)

-   details: link to further details (non-functional) (Click here for details)

```{r}
#| label - data description
#| dimension: ((17_186, 6), (2_065, 6))
radio_station <- tidytuesdayR::tt_load('2022-11-08')

unique(radio_station$station_info$service)
length(unique(radio_station$station_info$facility_id))
unique(radio_station$station_info$status)
unique(radio_station$station_info$details)

```

We selected the radio stations dataset because we thought it would be interesting to examine the relationships between radio stations and their region. We were especially interested in examining how the location of a radio station may be related to the category of its broadcast (called "format" in this dataset). For example, one thing we were curious about was which regions religious-leaning radio stations are the most concentrated. On its own, this dataset doesn't have many variables to work with. However, we plan to bring in external data with population statistics and demographic information for regions so we can conduct more analysis on how these factors may affect the characteristics of a radio station. 

# Questions

The two questions you want to answer.

# Analysis plan

A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
